For this lab, you will be joining and filtering related data sets to solve a murder mystery!

# Part 1: GitHub Workflow

At the top of the document (in the YAML) there is an `author` line that says `"Your name here!"`. Change this to be your name and save your file either by clicking on the blue floppy disk or with a shortcut (command / control + s).

Be sure to [commit the files to your repo](https://happygitwithr.com/existing-github-first#stage-and-commit).

Let's get started!

# Part 2: Some Words of Advice

-   Set chunk options carefully.

-   Make sure you don't print out more output than you need.

-   Make sure you don't assign more objects than necessary---avoid "object junk" in your environment.

-   Make your code readable and nicely formatted.

-   Think through your desired result **before** writing any code.

# Part 3: Finding the Killer

Northwestern University's Knight Lab wanted to help sharpen users' database skills, so they created a murder mystery. Can you solve this crime in SQL City??

The relational data you will be working with contains tables with different pieces of information pertinent to the crime - people, social media check-ins, driver's licenses, crime scene reports, police interviews, and more!

## Access the Data

This code chunk will read in **all** of the tables of data for you. Don't modify or remove this! Take some time to look at each file type so that

```{r}
#| author: Jacob Organista
#| message: false
#| warning: false
#| label: read-in-data

library(tidyverse)

# If purrr is not detected, install the package
if (!"purrr" %in% installed.packages()) install.packages("purrr")

source("https://raw.githubusercontent.com/jcanner/stat_210_2025_website/main/labs/instructions/lab-7-setup.R")
```

## Solve the Crime

### Crime Scene Report

Detective Wickham reaches out to you...

> A crime has taken place and I need your help! There was a murder in SQL City sometime on January 15, 2018. Could you retrieve the crime scene report from the police department’s database and follow the clues to find the person responsible for the murder?!

**Step 1: Find the police report in `crime_scene_report`. Then used the information in the report to move on to the next data source based on the information learned.**

```{r}
#| label: Filter reports
# Code for looking at the relevant crime scene report.
crime_scene_filter <- crime_scene_report %>%
  filter(
    date == 20180115, #date indicated
    city == "SQL City", #where was the report
    type == "murder" #what type of crime
  )

crime_scene_filter$description

```

**Next Steps: Follow the evidence to the person responsible for the murder, building a report as you go.** There are accomplices, some knowingly and some unknowingly, but there is only one mastermind.

Solve the murder mystery, showing **all of your work in this document**. Your document and code must be well organized, easy to follow, and reproducible.

-   Use headers and written descriptions to indicate what you are doing.
-   Use `dplyr` verbs and `join` functions rather than just looking through the tables manually. Functions from `stringr` and `lubridate` will also be useful.
-   Use good code formatting practices.
-   Comment your code.
-   Cite any external sources you use to solve the mystery.

```{r}
#| label: Filter witness
witness1 <- person %>%
  filter(address_street_name == "Northwestern Dr") %>% #street address of witness 1
  arrange(desc(address_number)) %>% #arranging to know who lives at the end
  slice(1) #only keeping the top to see witness

witness2 <- person %>%
  filter(address_street_name == "Franklin Ave", #street address of witness 2
         str_detect(name, "Annabel")) #looking for names containing _______

head(witness1)
head(witness2)
```

Notes:

-   Witness #1 Information:

    -   Name: Morty Schapiro

    -   License_id: 118009

    -   SSN: 111564949

    -   id: 14887

-   Witness #2 Information:

    -   Name: Annabel Miller

    -   License_id: 490173

    -   SSN: 318771143

    -   id: 16371

```{r}
#| label: witness transcripts
witness_ids <- bind_rows(witness1, witness2) %>% pull(id) #binding witness ids into one dataset

witness_interviews <- interview %>%
  filter(person_id %in% witness_ids) #getting the interviews of the witnesses
witness_interviews$transcript
```

New Information:

-   Man with "Get Fit Now Gym" bag (only gold members)

-   Membership start with: 48Z

-   Car Plate included: H42W

-   Recognized killer working out on January 9th

```{r}
#| label: Combining Suspect Information

possible_suspects <- drivers_license %>%
  filter(str_detect(plate_number, "H42W")) %>% #detecting the partial license plate numbers
  rename(license_id = id) %>% #renaming to avoid possible later confusion

  inner_join(person, by = "license_id") %>% #joinging person to drivers_license

  left_join(get_fit_now_member, by = c("id" = "person_id")) %>% #joining member df
  filter(membership_status == "gold") #filtering for the gold membership status

possible_suspects$name.x
```

```{r}
#| label: Matching Check-ins
combined_checkins <- bind_rows(
  get_fit_now_check_in %>%
    filter(membership_id == "48Z55"), #checking id to week prior to murder
  
  facebook_event_checkin %>% #checking for event check-ins that could be at time of murder
    filter(person_id == 67318)
)
head(combined_checkins)
```

Notes:

-   Jeremy was seen by Annabel a week ago on 2018/01/09

-   A week later the murder was committed by a gold member seen driving off with a partial plate (mentioned earlier) and partially matching membership number

-   Also a week later, Jeremy attended an event at the gym on 2018/01/15

    -   He was checked in on facebook, but not through the gym

```{r}
#| label: Suspect Interview
Jeremy_Interview <- interview %>%
  filter(
    person_id == 67318 #interviewing the supposed killer for hire?
  )
Jeremy_Interview$transcript
```

```{r}
#| label: Who Hired the Hitman?
final_suspect <- drivers_license %>%
  filter(
    height %in% c(65,67), #filter for guessed height
    hair_color == "red", #filter for hair color
    car_make == "Tesla", #filter for car make
    car_model == "Model S" #filter for model car
  ) %>%
 rename(license_id = id) %>% #renaming to match later dataframe

 inner_join(person, by = "license_id") %>% #joining person df to drivers_license

 inner_join(income, by = "ssn") #joining income df to drivers_license

final_suspect$name #name of final suspect
```

::: callout-note
## And the final suspect is...

Red Korb
:::
